---
title: "Data manipulation with dplyr"
author: "Ziqing Wang"
date: "2022-09-22"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import data 
```{r}
library(tidyverse)
options(tibble.print_min = 3)

litters_data = read_csv("./data/FAS_litters.csv", col_types = "ccddiiii")
litters_data = janitor::clean_names(litters_data)

pups_data = read_csv("./data/FAS_pups.csv")
pups_data = janitor::clean_names(pups_data)
```
# select

You can specify the columns you want to keep by naming all of them:

```{r}
select(litters_data, group, litter_number, gd0_weight, pups_born_alive)
```

You can specify the specify a range of columns to keep:

```{r}
select(litters_data, group:gd_of_birth) # from group variable to gd_of_birth
```
You can also specify columns you’d like to remove:

```{r}
select(litters_data, -pups_survive) # put a minus sign before the variable you want to remove
```
You can rename variables as part of this process.

```{r}
select(litters_data, GROUP = group, LiTtEr_NuMbEr = litter_number)

```
If all you want to do is rename something, you can use rename instead of select. This will rename the variables you care about, and keep everything else:

```{r}
rename(litters_data, GROUP = group, LiTtEr_NuMbEr = litter_number) # new_name = old_name
```
Helper functions that can be used in select():

```{r}
select(litters_data, starts_with("gd"))
```
I also frequently use is everything(), which is handy for reorganizing columns without discarding anything:

```{r}
select(litters_data, litter_number, pups_survive, everything()) # view litter_number and pups_survive as the first two variables
```
relocate() does a similar thing (and is sort of like rename in that it’s handy but not critical):

```{r}
relocate(litters_data, litter_number, pups_survive)
```
# filter 

```{r}
filter(litters_data, group == "Con7" & gd_of_birth == 20)
```

# mutate

create a new variable using existing variables

```{r}
mutate(litters_data,
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group)
)
```
# arrange

arrange the rows in your data according to the values in one or more columns:

```{r}
arrange(litters_data, group, pups_born_alive)
arrange(litters_data, desc(pups_born_alive)) # sort by descending order using desc() helper function
```

# piping (%>%)

```{r}
litters_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)) %>% 
  drop_na(wt_gain)

litters_data
```

```{r}
litters_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names(dat = .) %>%
  select(.data = ., -pups_survive) %>%
  mutate(.data = .,
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)) %>% 
  drop_na(data = ., wt_gain)
```


```{r}
litters_data %>%
  lm(wt_gain ~ pups_born_alive, data = .) %>% # placeholder
  broom::tidy()

```


